What Can Stack Traces Tell us about the Use of Exceptions in Java?

### 2014-07-02 android-stacks.tex
Including content about CLU language in Section2

The main disadvantage of return codes is the fact that the programmer can just
ignore the return code and let the program execute in inconsistent state. The
exception mechanism~\cite{goodenough1975exception} was proposed as a way of
preventing errors from being ignored, and was embedded in many mainstream
programming languages. Among some similarities they mainly differ in the way
exceptions are represented and handled in each
language~\cite{garcia2001comparative}:

\noindent\emph{Checked Approach}. In languages such Modula-3, Guide, and
Extended Ada all exceptions are checked, which means that the compiler
statically checks if appropriate handlers are provided within the system.
Accordingly, the programmer must explicitly specify every exception that can be
signaled to support the compile time checking. This approach is known to be
costly to maintain (Dooren and Steegmans, 2005) since exception updates are
cascading; if a new exception type is added to a method signature then all other
methods directly using this method must either change their signatures to
include the new exception type or handle the new exception.

\noindent\emph{Unchecked Approach}. In languages such as Lore, C++, and Arche
all exceptions are unchecked but the developer can optionally list exception
types on method signatures. However, the developer is not warned by the compiler
if an unchecked exception is not handled inside the application. Such an
approach may hinder robustness because the client of a method cannot easily know
which unchecked exceptions may be thrown, unless the code of the method and the
methods called from it is inspected, which can be a very time consuming or
infeasible task.

\noindent\emph{Hybrid Approach}. In Java a method can throw either checked or
unchecked exceptions. Java is the only language that allows this hybrid
approach.

@@

Table~\ref{tab:toptenjava} illustrates the top 10 most common root causes which
represents more than 50\% of the analyzed set. The table illustrates  i) the
number of times each exception appeared as the root cause of a stack trace
(\#occurrences), ii) the number of distinct projects reporting such causes on
stack traces, and iii) the exception popularity among projects.

>>>>

The main disadvantage of return codes is the fact that the programmer can just
ignore the return code and let the program execute in inconsistent state. The
exception mechanism~\cite{goodenough1975exception} was proposed as a way of
preventing errors from being ignored and was embedded in many mainstream
programming languages ~\cite{garcia2001comparative} since then.

\noindent\emph{Checked Approach}. In languages such as CLU \footnote{In 
CLU if a handler is found for a given exception, it is converted in a predened 
default exception, called failure which is the only one which can be implicitly 
propagated in this model.}, Modula-3, Guide, and Extended Ada all exceptions 
are checked, which means that the compiler
statically checks if appropriate handlers are provided within the system.
Accordingly, the programmer must explicitly specify every exception that can be
signaled to support the compile time checking. This approach is known to be
costly to maintain (Dooren and Steegmans, 2005) since exception updates are
cascading; if a new exception type is added to a method signature then all other
methods directly using this method must either change their signatures to
include the new exception type or handle the new exception.

\noindent\emph{Unchecked Approach}. In languages such as Lore, C++, and Arche
all exceptions are unchecked but the developer can optionally list exception
types on method signatures. Although in some languages such exceptions can 
be documented on the methods signature, the developer is not warned by the compiler
if an unchecked exception is not handled inside the application. According to ~\cite{Robil00} 
such an approach may hinder robustness because the client of a method cannot easily know
which unchecked exceptions may be thrown (unless the code of the method and the
methods called from it is inspected) which can be a very time consuming or
infeasible task.

\noindent\emph{Hybrid Approach}. In Java a method can throw either checked or
unchecked exceptions. Java is the only language that allows this hybrid approach,
on which the developer can define exceptions of both types of exceptions to be thrown and
handled in a single system.

@@

Table~\ref{tab:toptenjava} illustrates the top 10 most common root causes which
represents more than 50\% of the analyzed set. The table illustrates  i) the
number of times each exception appeared as the root cause reported on the exception stack trace
(\#occurrences), ii) the number of distinct projects reporting such causes on
stack traces, and iii) the exception popularity among projects.
